---
title: "Alveolar Macrophages Pathway Enrichment Analysis"
author: "Iliana Papadopoulou"
date: "`r Sys.Date()`"
output:
  html_document:
    self_contained: yes
    code_folding: show
    number_sections: no
    toc: yes
    toc_depth: 3
    toc_float: true
    collapsed: no
    smooth_scroll: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE, dev = c("png", "pdf"))
set.seed(1000)
options(future.globals.maxSize = 8000 * 1024^2)
```

## Customize parameters

```{r Customization}
library(plyr)
library(biomaRt)
library(dplyr)
library(tidyverse)
library(simplifyEnrichment)
library(msigdbr)
library(msigdbr)
library(here)
library(stats)
library(ggsci)
library(clusterProfiler)
library(dict)
library(org.Mm.eg.db)
library(DOSE)
library(Seurat)
library(openxlsx)
library(zeallot)
library(ggsci)
library(openxlsx)
library(pheatmap)
library(stringr)
library(viridis)
library(scCustomize)
library(stringr)
library(gridExtra)
library(grid)

httr::set_config(httr::config(ssl_verifypeer = FALSE))

organism <- "org.Mm.eg.db"
disease_phase <- ""
datatype <- 'SYMBOL'

```

# **4.Pathway enrichment analysis of control vs PH-cHFpEF mice in alveolar macrophages**

## Set directory
```{r set directory}
basedir <- here()
source(paste0(basedir,'/analysis/Lung_snRNAseq_functions.R'), echo=FALSE)

# Create directory to save preprocessed seurat objects
output_obj_dir <- paste0(basedir,"/data/Output/")
dir.create(output_obj_dir)
```

## Create directory for saving clustering results
```{r directory doe clustering results}
GSEADir <- paste0(output_obj_dir,"GSEA")
dir.create(GSEADir)
```

## Read seurat annotated object
```{r read data}
LungMu_obj <- readRDS(file= paste0(output_obj_dir, "340821_Mu_Lung_snRNAseq_merged_annotated.rds" ))

LungMu_obj$celltype_condition <- paste0(LungMu_obj$label,"_",LungMu_obj$condition)
```

## Color vectors - Define color palete
```{r color palet}
colCond <- c("#631879FF","#F16913")
names(colCond) <- unique(LungMu_obj$condition)
```


```{r GSEA analysis}
GSEA_Analysis <- function(data,organism,datatype,disease_phase,celltype,outdir){
  obj_sub <- subset(data, label == celltype)
  print(unique(obj_sub$label))
  Idents(obj_sub) <- obj_sub$condition
  markers <- FindAllMarkers(object = obj_sub, 
                            only.pos = TRUE,
                            logfc.threshold = 0.25,
                            min.pct = 0.1,
                            test.use = "wilcox")  
  
  Vec <- sort(unique(obj_sub$condition))
  EnrichParameters <-customize_parameters(Vec,markers,organism,datatype,disease_phase, GSEADir)
  
  for (i in seq(1,length(EnrichParameters$enrichcl_list))){
    terms<- EnrichParameters$enrichcl_list[[i]]@result
    condition <- Vec[i]
    condition <- gsub("/", "_", condition)
    celltype <- gsub("/", "_", celltype)
    write.xlsx(terms, paste0(outdir,"/","GO_Pathways_",celltype,"_",condition,".xlsx"),rowNames = TRUE)
  } 
  return(EnrichParameters)
}


celltype <-"Alveolar Mph"
data <- LungMu_obj
EnrichParameters <- GSEA_Analysis(data,organism,datatype,disease_phase,celltype,GSEADir)
```

## Join files in order to gather all info in one file - for gene name
```{r join pathway files}
# For control
GO_semantic_Al_Mph_CD <- read.xlsx(paste0(GSEADir,"/","GO_Pathways_Alveolar Mph_Control.xlsx"))
GO_semantic_Al_Mph_CD[,1] <- NULL

# For PH-cHFpEF
GO_semantic_HFD <- read.xlsx(paste0(GSEADir,"/","GO_Pathways_Alveolar Mph_PH-cHFpEF.xlsx"))
GO_semantic_HFD[,1] <- NULL
```

## Find overlapping pathways between Alveolar macrophages control_vs_PH-cHFpEF vs control
```{r compute overlapping pathways}
df_overlapping_pathways <- inner_join(GO_semantic_Al_Mph_CD, GO_semantic_HFD,by=c("ID","Description") )
overlapping_pathways <- df_overlapping_pathways$ID
```

## Find exclusive pathways in both  Alveolar macrophages control_vs_PH-cHFpEF vs control
```{r compute exclusive pathways}
df_list <-list(GO_semantic_Al_Mph_CD, GO_semantic_HFD)
GO_semantic_AlvMph<-do.call("rbind", df_list)
GO_semantic_AlvMph <- GO_semantic_AlvMph %>%
  filter (! ID %in% overlapping_pathways)
```

## Figure 2E - Plot selected overlapping pathways and exclusive between Alveolar macrophages control vs PH-cHFpEF mice
```{r visualite pathways, fig.width=8}
pathways <- c(
"GO:0050881","GO:0086001",  "GO:0098760", "GO:0098779",
"GO:0043277","GO:0031589","GO:0030509","GO:1903036", "GO:0071353","GO:2000343","GO:0071248"
)


t_df_overlapping_pathways <- df_overlapping_pathways
PATH_per_cond <- t_df_overlapping_pathways[,c("ID","Description", "p.adjust.x", "analysis.x", "p.adjust.y", "analysis.y")]

# For pathways that are shared between conditions
PATH_per_cond <- PATH_per_cond %>% filter(ID %in% pathways, .keep_all = TRUE)

rownames(PATH_per_cond) <- PATH_per_cond$Description
PATH_per_cond$Description <- NULL
PATH_per_cond$ID <- NULL

colnames(PATH_per_cond) <- c(unique(PATH_per_cond$analysis.x), 
                             "analysis.x",
                            unique(PATH_per_cond$analysis.y),
                             "analysis.y")
PATH_per_cond$analysis.x <- NULL
PATH_per_cond$analysis.y <- NULL

# For pathways that are exclusive in each condition
exl_path <- setdiff(pathways, PATH_per_cond$ID)
exl_path_cond <- GO_semantic_AlvMph[,c("ID","Description", "p.adjust", "analysis")]
exl_path_cond <- exl_path_cond %>% filter(ID %in% exl_path, .keep_all = TRUE)
exl_path_cond <- exl_path_cond %>%
  pivot_wider(id_cols = c('ID','Description'), names_from = 'analysis',
              values_from = 'p.adjust') %>%
  as.data.frame()
rownames(exl_path_cond) <- exl_path_cond$Description
exl_path_cond$Description <- NULL
exl_path_cond$ID <- NULL

 if (("Control" %in% colnames(exl_path_cond) == "FALSE")){
   exl_path_cond$Control <- NA
 }

# Combine two dfs
PATH_per_cond <- rbind(PATH_per_cond, exl_path_cond)

# Log the dataframe
PATH_per_cond[] <- lapply(PATH_per_cond, function(x) -1 *(log10(x)))
PATH_per_cond[is.na(PATH_per_cond)] <- 0

PATH_per_cond <- PATH_per_cond %>% 
  mutate(across(where(is.numeric))/rowSums(across(where(is.numeric))))

# Generate annotations for columns
annotation_col = data.frame(
                    condition = colnames(PATH_per_cond)
                )
rownames(annotation_col) = colnames(PATH_per_cond)
ann_colors = list(condition = colCond)

# Choose color palette
palette_length = 100
     
pheatmap(PATH_per_cond, cluster_rows=T, cluster_cols=F, cellwidth = 10, cellheight = 10, scale = "none",
        color = colorRampPalette(RColorBrewer::brewer.pal(n = 11, name = "RdYlBu") )(50)|> rev() ,
        annotation_col = annotation_col, annotation_colors = ann_colors) 
```

## Session Information
````{r Session Information}
sessionInfo()
````