---
title: "Alveolar Macrophages -  Endothelial Cells Interactome analysis"
author: "Iliana Papadopoulou"
date: "`r Sys.Date()`"
output:
  html_document:
    self_contained: yes
    code_folding: show
    number_sections: no
    toc: yes
    toc_depth: 3
    toc_float: true
    collapsed: no
    smooth_scroll: yes
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE, dev = c("png", "pdf"))
set.seed(1000)
options(future.globals.maxSize = 8000 * 1024^2)
```

## Load packages
```{r libraries}
suppressPackageStartupMessages({
  library(plyr)
  library(patchwork)
  library(Seurat)
  library(stringr)
  library(dplyr)
  library(tidyverse)
  library(here)
  library(openxlsx)
  library(zeallot)
  library(ggsci)
  library(openxlsx)
  library(pheatmap)
  library(stringr)
  library(viridis)
  library(scCustomize)
  library(stringr)
  library(gridExtra)
  library(grid)
  library(CellChat)
  library(ComplexHeatmap)
})
```


# **5. Interactome analysis among alveolar macrophages and endothelial cells in control vs PH-cHFpEF mice**

## Set directory
```{r set directory}
basedir <- here()
source(paste0(basedir,'/analysis/Lung_snRNAseq_functions.R'), echo=FALSE)
source(paste0(basedir,'/analysis/Cellchat_fun.R'), echo=FALSE)
# Create directory to save the analysis outcome


# Create directory to save preprocessed seurat objects
output_obj_dir <- paste0(basedir,"/data/Output/")
dir.create(output_obj_dir)
```

## Read seurat annotated object
```{r read object}
LungMu_obj <- readRDS(file= paste0(output_obj_dir, "340821_Mu_Lung_snRNAseq_merged_annotated.rds" ))
```

## Define color palette
```{r color palette}
colCond <- c("#631879FF","#FF7F0EFF")
names(colCond) <- unique(LungMu_obj$condition)
```

## If object contains ensemble in gene name, remove it
```{r remove ENSG id}

#Convert to SingleCellExperimentObject for changing the rownames without change the reductions
LungMu_obj.sce <- as.SingleCellExperiment(LungMu_obj, assay = "RNA") 

gene_name_fixdf <- data.frame(Gene_Name = rownames(LungMu_obj.sce))
gene_name_fixdf$Symbol <- gsub("^[^.]*.","", gene_name_fixdf$Gene_Name)

# Check duplicates in df and fix them
duplidx <- rownames((gene_name_fixdf[duplicated(gene_name_fixdf$Symbol), ]))
Fixed_dupl <- gene_name_fixdf %>% 
  filter(row_number() %in% duplidx)  %>% 
 mutate(Symbol = paste0(Symbol, ".1"))
gene_name_fixdf[duplidx,"Symbol"] <- Fixed_dupl$Symbol

# Check again for double duplicates
duplidx <- rownames((gene_name_fixdf[duplicated(gene_name_fixdf$Symbol), ]))

Fixed_dupl <- gene_name_fixdf %>% 
  filter(row_number() %in% duplidx)  %>% 
 mutate(Symbol = paste0(Symbol, ".2"))
gene_name_fixdf[duplidx,"Symbol"] <- Fixed_dupl$Symbol

rownames(LungMu_obj.sce) <- gene_name_fixdf$Symbol

# Convert SingleCell experiment object back to seurat object by retaining the reductions, normalization, etc
LungMu_obj <- as.Seurat(LungMu_obj.sce, assay = NULL)
```

## Select alveolar macrophages and endothelial cells
```{r cell population selection}
celltypes_involved <- c("aCap pulmonary EC", "Arterial/Cap pulmonary EC", "Alveolar Mph")
LungMu_obj <- subset(LungMu_obj, label %in% celltypes_involved)
LungMu_obj_CD <- subset(LungMu_obj, condition == "Control")
LungMu_obj_HFD <- subset(LungMu_obj, condition == "PH-cHFpEF")
```

## Cell type labels
```{r}
col_annotated <- pal_igv()(length(unique(LungMu_obj$label)))
names(col_annotated) <- unique(LungMu_obj$label)

col_annotated["Arterial/Cap pulmonary EC"] <-"#990080FF"
col_annotated["Alveolar Mph"] <-"#00CC33FF"
col_annotated["aCap pulmonary EC"] <- "#0099CCFF"
```

## Interactome Analysis
```{r}
## read object into cellchat
cellchat_CD <- createCellChat(LungMu_obj_CD, group.by = "label")
cellchat_HFD <- createCellChat(LungMu_obj_HFD, group.by = "label")

cellchat_CD <- CellToCellCommunicationAnalysis(cellchat_CD,"mouse")

## Save cellChat object for faster processing
saveRDS(cellchat_CD, file= paste0(output_obj_dir, "340821_Mu_snRNAseq_lung_cellchat_Control_Control_vs_PH-cHFpEF_AlvMph_EC.rds"))

cellchat_HFD <- CellToCellCommunicationAnalysis(cellchat_HFD,"mouse")

## Save cellChat object for faster processing
saveRDS(cellchat_HFD, file= paste0(output_obj_dir, "340821_Mu_snRNAseq_lung_cellchat_PH-cHFpEF_control_vs_PH-cHFpEF_AlvMph_EC.rds"))
```

## Merge cellchat objects
```{r merge objects}
object.list <- list(Control = cellchat_CD, PHcHFpEF = cellchat_HFD)
cellchat <- mergeCellChat(object.list, add.names = names(object.list))
```

## Compare the total number of interactions and interaction strength --> Communication is enhanced in control vs FPEF
```{r number and strength of interactions}
gg1 <- compareInteractions(cellchat, show.legend = F, group = c(1,2), color.use = unname(colCond))
gg2 <- compareInteractions(cellchat, show.legend = F, group = c(1,2), measure = "weight",color.use = unname(colCond))
gg1 + gg2
```

## Compare the major sources and targets in 2D space
### Identify cell populations with significant changes in sending or receiving signals
```{r}
num.link <- sapply(object.list, function(x) {rowSums(x@net$count) + colSums(x@net$count)-diag(x@net$count)})
weight.MinMax <- c(min(num.link), max(num.link)) # control the dot size in the different datasets
gg <- list()
for (i in 1:length(object.list)) {
  gg[[i]] <- netAnalysis_signalingRole_scatter(object.list[[i]], title = names(object.list)[i], weight.MinMax = weight.MinMax, color.use = col_annotated)
}
patchwork::wrap_plots(plots = gg)
```


## Compare the overall information flow of each signaling pathway to identify altered signaling with distinct interaction strength
```{r}
gg1 <- rankNet(cellchat, mode = "comparison", measure = "weight", sources.use = NULL, targets.use = NULL, stacked = T, do.stat = TRUE, color.use = unname(colCond))
gg1
```

## Figure 7A : Compute signaling pathways exchanged among endothelial cells and alveolar macrophages via heatmaps{.tabset}
### Control mice
```{r Control mice}
col_annotated_reorder <- col_annotated[order(factor(names(col_annotated), 
                                             levels = rownames(object.list[[1]]@netP$prob)))]

col_annotated_reorder <- col_annotated_reorder[names(col_annotated_reorder) %in% celltypes_involved]
i = 1
# combining all the identified signaling pathways from different datasets 
pathway.union <- union(object.list[[i]]@netP$pathways, object.list[[i+1]]@netP$pathways)
ht1 = netAnalysis_signalingRole_heatmap(object.list[[i]], pattern = "outgoing", signaling = pathway.union, title = names(object.list)[i], width = 5, height = 5, color.heatmap = "YlOrRd", color.use = col_annotated_reorder)
ht2 = netAnalysis_signalingRole_heatmap(object.list[[i+1]], pattern = "outgoing", signaling = pathway.union, title = names(object.list)[i+1], width = 5, height = 5, color.heatmap = "YlOrRd", color.use = col_annotated_reorder)

i = 1
# combining all the identified signaling pathways from different datasets 
pathway.union <- union(object.list[[i]]@netP$pathways, object.list[[i+1]]@netP$pathways)
ht3 = netAnalysis_signalingRole_heatmap(object.list[[i]], pattern = "incoming", signaling = pathway.union, title = names(object.list)[i], width = 5, height = 5, color.heatmap = "YlOrRd", color.use = col_annotated_reorder)
ht4 = netAnalysis_signalingRole_heatmap(object.list[[i+1]], pattern = "incoming", signaling = pathway.union, title = names(object.list)[i+1], width = 5, height = 5, color.heatmap = "YlOrRd", color.use = col_annotated_reorder)

draw(ht1 + ht3, ht_gap = unit(0.5, "cm"))
```

### PH-cHFpEF mice
```{r PHcHFpEF}
draw(ht2 + ht4, ht_gap = unit(0.5, "cm"))
```


## Ligand-receptor (LR) interaction pairs{.tabset}
### Exchanged signaling Endothelial - Alveolar Macrophages
```{r EC-Alv Mph}
celltypes_involved <- c("aCap pulmonary EC", "Arterial/Cap pulmonary EC", "Alveolar Mph")
cols_vector <- col_annotated[names(col_annotated) %in% celltypes_involved ]

## EC --> Alveolar Mph
for (i in 1:length(object.list)) {
  netVisual_chord_gene(object.list[[i]], sources.use = c(1,3) , targets.use = c(2), lab.cex = 0.8, color.use=col_annotated_reorder ,title.name = paste0("EC to Alveolar Mph  - ", names(object.list)[i]), small.gap = 1,show.legend = TRUE,legend.pos.x = 4,legend.pos.y = 5)
}
```

### Exchanged signaling Alveolar Macrophages - Endothelial
```{r Alv Mph-EC}
netVisual_chord_gene(object.list[[1]], sources.use = c(2) , targets.use = c(1,3), lab.cex = 0.9, color.use=col_annotated_reorder ,title.name = paste0("Alveolar Mph to EC  - ", names(object.list)[1]), small.gap = 1,show.legend = TRUE,legend.pos.x = 8,
                     legend.pos.y = 7)

```

## Session Information
````{r Session Information}
sessionInfo()
````
