---
title: "MuLung_snRNAseq_analysis"
author: "Iliana Papadopoulou"
date: "`r Sys.Date()`"
output:
  html_document:
    self_contained: yes
    code_folding: show
    number_sections: no
    toc: yes
    toc_depth: 3
    toc_float: true
    collapsed: no
    smooth_scroll: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE, message = FALSE, dev = c("png", "pdf"))
set.seed(1000)
options(future.globals.maxSize = 8000 * 1024^2)
```

# 2. Analysis of murine lung samples from healthy mice vs mice with HFpEF in Pulmonary Hypertension (PH)

## Load R libraries
```{r libraries}
suppressPackageStartupMessages({
  library(plyr)
  library(patchwork)
  library(Seurat)
  library(stringr)
  library(dplyr)
  library(tidyverse)
  library(here)
  library(openxlsx)
  library(zeallot)
  library(ggsci)
  library(openxlsx)
  library(pheatmap)
  library(stringr)
  library(viridis)
  library(scCustomize)
  library(stringr)
  library(gridExtra)
  library(grid)
})
```

## Set directory
```{r set directory}
basedir <- here()
source(paste0(basedir,'/analysis/Lung_snRNAseq_functions.R'), echo=FALSE)

# Create directory to save preprocessed seurat objects
output_obj_dir <- paste0(basedir,"/data/Output/")
dir.create(output_obj_dir)
```

## Read preprocessed murine lung seurat object
```{r}
LungMu_obj <- readRDS(file= paste0(output_obj_dir, "340821_Mu_Lung_snRNAseq_merged_preprocessed.rds" ))
```

## Color vectors - Define color palete
```{r color palete}
cols<- pal_igv()(51)
names(cols) <- c(0:50)

colDataset <- c("#FF7F0EFF","#17BECFFF", "#BB0021FF", "#00CC33FF", "#3B4992FF", "#4775FFFF", "#990080FF", "#AEC7E8FF","#FFBB78FF","#1B1919FF")
names(colDataset) <- unique(LungMu_obj$dataset)

colCond <- c("#631879FF","#FF7F0EFF") #"#BCBD22FF"
names(colCond) <- unique(LungMu_obj$condition)
```


## Preprocessing
```{r preprocessing}
resolution <- c(0.1, 0.15,0.2, 0.25,0.3, 0.4, 0.5, 0.6, 0.8, 1., 1.2, 1.5, 1.7, 2, 2.2, 2.4, 2.6)
LungMu_obj  <- preprocessing(LungMu_obj,resolution)
```

## Cell type annotation
```{r annotation}
LungMu_obj$label <- "other"
LungMu_obj$label[which(LungMu_obj$RNA_snn_res.0.15 %in% c("0"))] <- "AT2 Epithelial"
LungMu_obj$label[which(LungMu_obj$RNA_snn_res.0.15 %in% c("1"))] <-  "AT1 Epithelial"
LungMu_obj$label[which(LungMu_obj$RNA_snn_res.0.15 %in% c("2"))] <- "Bas/SMG duct Epithelial 1"
LungMu_obj$label[which(LungMu_obj$RNA_snn_res.0.15 %in% c("3"))] <- "Fibroblasts"
LungMu_obj$label[which(LungMu_obj$RNA_snn_res.0.15 %in% c("4"))] <- "Arterial/Cap pulmonary EC"
LungMu_obj$label[which(LungMu_obj$RNA_snn_res.0.15 %in% c("5"))] <- "Alveolar Mph"
LungMu_obj$label[which(LungMu_obj$RNA_snn_res.0.15 %in% c("6"))] <- "Bas/SMG duct Epithelial 2"
LungMu_obj$label[which(LungMu_obj$RNA_snn_res.0.15 %in% c("7"))] <- "Cil Epithelial"
LungMu_obj$label[which(LungMu_obj$RNA_snn_res.0.15 %in% c("8"))] <-  "Immune cells"
LungMu_obj$label[which(LungMu_obj$RNA_snn_res.0.15 %in% c("9"))] <- "aCap pulmonary EC"
LungMu_obj$label[which(LungMu_obj$RNA_snn_res.0.15 %in% c("10"))] <-  "Mesothelial"
LungMu_obj$label[which(LungMu_obj$RNA_snn_res.0.15 %in% c("11"))] <-  "VSMC/Pericytes"

colLab <- pal_igv()(length(unique(LungMu_obj$label)))
names(colLab) <- unique(LungMu_obj$label)

colLab["AT2 Epithelial"] <-"#92C5DE"
colLab["Bas/SMG duct Epithelial 1"] <-"#F4A582" 
colLab["Bas/SMG duct Epithelial 2"] <-"#FDDBC7" 
colLab["AT1 Epithelial"] <-"#9467BDFF"
colLab["Fibroblasts"] <- "#FF1463FF"
colLab["Arterial/Cap pulmonary EC"] <-"#990080FF"
colLab["Alveolar Mph"] <-"#00CC33FF"
colLab["Cil Epithelial"] <- "#009966FF"
colLab["Immune cells"] <- "#053061"
colLab["aCap pulmonary EC"] <- "#0099CCFF"
colLab["Mesothelial"] <- "#67001F"
colLab["VSMC/Pericytes"] <- "#E7C76FFF"

```

## Sample Visualization and Clustering {.tabset}
### Figure 2A - cluster annotation
```{r Figure 2A - cluster annotation}

colLab_reorder <- colLab[order(factor(names(colLab), levels = c("Arterial/Cap pulmonary EC","aCap pulmonary EC",
                                                                "Bas/SMG duct Epithelial 1","Bas/SMG duct Epithelial 2","AT1 Epithelial","AT2 Epithelial", "Cil Epithelial",
                                                                "Fibroblasts","VSMC/Pericytes","Mesothelial",
                                                                "Alveolar Mph", "Immune cells")))]

DimPlot(LungMu_obj, reduction = "umap", group.by = "label", cols = colLab_reorder, order =  names(rev(colLab_reorder)))+
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text = element_blank(), axis.ticks = element_blank(), 
    panel.grid.minor = element_blank()) +
  ggtitle("") +
  xlab("UMAP1") +
  ylab("UMAP2")

```

### Figure 2B - Umap per condition
```{r Umap per condition}
DimPlot(LungMu_obj, reduction = "umap", group.by = paste0("condition"), cols = colCond)+
  theme_bw() +
  theme( axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    panel.grid.minor = element_blank()) +
  ggtitle("") +
  xlab("UMAP1") +
  ylab("UMAP2")

```

### Umap per dataset
```{r Umap per dataset, fig.width=9}
DimPlot(LungMu_obj, reduction = "umap", group.by = paste0("dataset"), cols = colDataset)+
  theme_bw() +
  theme( axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text = element_blank(), axis.ticks = element_blank(), 
    panel.grid.minor = element_blank()) +
  ggtitle("") +
  xlab("UMAP1") +
  ylab("UMAP2")
```  

## Compute DE genes for all clusters
```{r compute DE genes for all clusters}
Idents(LungMu_obj) <- "label"
de.markers <- FindAllMarkers(LungMu_obj, logfc.threshold = 0.25,min.pct = 0.1,only.pos = TRUE)

# plot DE genes top 10 avg logFC
analysis_type <- "total_marker_genes"

genes <- data.frame(gene=rownames(LungMu_obj)) %>% 
    mutate(geneID=gsub("^.*\\.", "", gene)) 

Selgenes <- c("Hey1", "Cxcl12", "Notch4", "Edn1",
              "Ednrb",
              "Muc5b","Krt19","Dlk2", 
              "Klf5","Hes1","Mgst1", "Klf15",
              "Clic5", "Spock2", "Ager", "Hopx", 
             "Sftpb", "Sftpa1", "Etv5", "Acsl4",
              "Foxj1", "Caps2", "Ccdc151", "Dlec1",
             "Col1a2" , "Col1a1", "Pdgfra",
             "Myh11", "Tagln", "Pdgfrb", "Carmn",
              "Muc16", "Msln", "Wt1", "Bnc1",
              "Itgax", "Marco", "Siglecf","Chil3",
              "F13a1","Ighm", "Il7r", "Klrd1"
              )

Selgenes_df <- filter(genes, geneID %in% Selgenes)
colnames(Selgenes_df) <- c('geneID','gene')
# order by Selgenes
Selgenes_df <-Selgenes_df[match(Selgenes, Selgenes_df$gene),]
```

## Explore cell populations{.tabset}
### Figure S2A - Visualize DE genes over all cell populations
```{r Visualize DE genes over all cell populations, fig.height = 12}
Idents(LungMu_obj) <- LungMu_obj$label
pOut <- avgHeatmapTotalHeatmap(seurat = LungMu_obj, selGenes = Selgenes_df,
                  colVecIdent = colLab_reorder, 
                  ordVec=levels(LungMu_obj),
                  gapVecR=c(4,5,8, 12,16,20, 24, 27, 31, 35, 39 ), gapVecC=NULL,cc=F,
                  cr=F, condCol=F,
                  path = GeneExprDir,
                  analysis_type = analysis_type,
                  len = 0)


```

### Figure S2B - Differences per condition
```{r Differences per condition, fig.width= 9}
DimPlot(LungMu_obj,split.by = "condition",cols = colLab, group.by = "label") 
```

### Figure S3C - Plot distribution of Pulmonary cells in control mice
```{r distribution of Pulmonary cells in control  mice}
percentage_matrix  <- as.data.frame(LungMu_obj@meta.data %>% dplyr::count(condition, label) %>% dplyr::group_by(condition) %>% mutate(cell_percentage= prop.table(n) * 100)) # RNA_snn_res.0.2

distr_df <- percentage_matrix[percentage_matrix$condition == "Control",]
distr_df$condition <- NULL
names(distr_df)[names(distr_df) == 'label'] <- 'sample'
names(distr_df)[names(distr_df) == 'n'] <- 'cells'
names(distr_df)[names(distr_df) == 'cell_percentage'] <- 'fraction'
distr_df$cell_percentage <- distr_df$fraction / 100

cells_per_datasetDF <- distr_df
# Compute the cumulative percentages (top of each rectangle)
cells_per_datasetDF$ymax <- cumsum(cells_per_datasetDF$fraction)

# Compute the bottom of each rectangle
cells_per_datasetDF$ymin <- c(0, head(cells_per_datasetDF$ymax, n=-1))

# Compute label position
cells_per_datasetDF$labelPosition <- (cells_per_datasetDF$ymax + cells_per_datasetDF$ymin) / 2

# Compute a good label
cells_per_datasetDF$label <- paste0(cells_per_datasetDF$patient, "\n value: ", cells_per_datasetDF$nuclei)

# Make the plot
p<- ggplot(cells_per_datasetDF, aes(ymax=ymax, ymin=ymin, xmax=4, xmin=3, fill=sample)) +
  geom_rect() +
  scale_fill_manual(values = colLab_reorder) +
  coord_polar(theta="y", start = 0) +
  xlim(c(2, 4)) +
  theme_void() +
  annotate(geom = 'text', x = 2, y = 0, label = as.character(dim(subset(LungMu_obj,condition == "Control" ))[2]), size = 10)
p
```

### Figure S3C - Plot distribution of Pulmonary cells in HFD mice
```{r distribution of Pulmonary cells in HFD  mice}
percentage_matrix  <- as.data.frame(LungMu_obj@meta.data %>% dplyr::count(condition, label) %>% dplyr::group_by(condition) %>% mutate(cell_percentage= prop.table(n) * 100)) 

distr_df <- percentage_matrix[percentage_matrix$condition == "PH-cHFpEF",]
distr_df$condition <- NULL
names(distr_df)[names(distr_df) == 'label'] <- 'sample'
names(distr_df)[names(distr_df) == 'n'] <- 'cells'
names(distr_df)[names(distr_df) == 'cell_percentage'] <- 'fraction'
distr_df$cell_percentage <- distr_df$fraction / 100

cells_per_datasetDF <- distr_df
# Compute the cumulative percentages (top of each rectangle)
cells_per_datasetDF$ymax <- cumsum(cells_per_datasetDF$fraction)

# Compute the bottom of each rectangle
cells_per_datasetDF$ymin <- c(0, head(cells_per_datasetDF$ymax, n=-1))

# Compute label position
cells_per_datasetDF$labelPosition <- (cells_per_datasetDF$ymax + cells_per_datasetDF$ymin) / 2

# Compute a good label
cells_per_datasetDF$label <- paste0(cells_per_datasetDF$patient, "\n value: ", cells_per_datasetDF$nuclei)

# Make the plot
p<- ggplot(cells_per_datasetDF, aes(ymax=ymax, ymin=ymin, xmax=4, xmin=3, fill=sample)) +
  geom_rect() +
  scale_fill_manual(values = colLab_reorder) +
  coord_polar(theta="y", start = 0) +
  xlim(c(2, 4)) +
  theme_void() +
  annotate(geom = 'text', x = 2, y = 0, label = as.character(dim(subset(LungMu_obj,condition == "PH-cHFpEF" ))[2]), size = 10)
p
```

### Stacked Barplot : Cell type abundance per sample
```{r  Cell type abundance per sample, fig.height= 9}
# Run successfully
percentage_matrix  <- as.data.frame(LungMu_obj@meta.data %>% count(dataset, label) %>% group_by(dataset) %>% mutate(cell_percentage= prop.table(n) * 100))

# Fix order of celltypes
percentage_matrix$label  <- factor (percentage_matrix$label ,
          levels = names(colLab_reorder),
          ordered = TRUE)

p<- ggplot(percentage_matrix,    # Calculate percent within each dataset
       aes(dataset, cell_percentage, fill=label)) +
  geom_bar(stat="identity") + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        axis.line = element_line(), # colour = "black"
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank()) +
  geom_density() + # color = "black"
  labs(y= "Cell type (%)" ) +
  scale_fill_manual(values = colLab_reorder)
p
```

## Classify cell types into endothelial, epithelial, immune and stroma cells
```{r plot cells per endothelial, epithelial, imuune and stroma cells }
LungMu_obj$type <- "Other"
LungMu_obj$type[which(LungMu_obj$label %in% c("Arterial/Cap pulmonary EC", "aCap pulmonary EC"))] <- "Endothelial"
LungMu_obj$type[which(LungMu_obj$label %in% c("Bas/SMG duct Epithelial 1", "Bas/SMG duct Epithelial 2","AT2 Epithelial", "AT1 Epithelial", "Cil Epithelial"))] <- "Epithelial"
LungMu_obj$type[which(LungMu_obj$label %in% c("Immune cells", "Alveolar Mph"))] <- "Immune"
LungMu_obj$type[which(LungMu_obj$label %in% c("Fibroblasts", "Mesothelial", "VSMC/Pericytes"))] <- "Stroma"

colType <- c("#374E55FF", "#D6604D", "#79AF97FF", "#0099CCFF")
names(colType) <- unique(LungMu_obj$type)

DimPlot(LungMu_obj, reduction = "umap", group.by = "type", cols = colType)+
  theme_bw() +
  theme(axis.line = element_line(colour = "black"),
    panel.grid.major = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.text = element_blank(), axis.ticks = element_blank(), 
    panel.grid.minor = element_blank()) +
  ggtitle("") +
  xlab("UMAP1") +
  ylab("UMAP2")
```

## Find differences per condition
```{r}
for (cellType in unique(LungMu_obj$label)){
  LungMu_obj_sub <- subset(LungMu_obj, label == cellType)
  Idents(LungMu_obj_sub) <- "condition"
  de.markers <- FindAllMarkers(LungMu_obj_sub, logfc.threshold = 0.25,min.pct = 0.1,only.pos = TRUE)
}

LungMu_obj_sub <- subset(LungMu_obj, label == "Bas/SMG duct Epithelial 1")
Idents(LungMu_obj_sub) <- "condition"
de.markers <- FindAllMarkers(LungMu_obj_sub, logfc.threshold = 0.25,min.pct = 0.1,only.pos = TRUE)

LungMu_obj_sub <- subset(LungMu_obj, label == "Bas/SMG duct Epithelial 2")
Idents(LungMu_obj_sub) <- "condition"
de.markers <- FindAllMarkers(LungMu_obj_sub, logfc.threshold = 0.25,min.pct = 0.1,only.pos = TRUE)

LungMu_obj_sub <- subset(LungMu_obj, label == "VSMC/Pericytes")
Idents(LungMu_obj_sub) <- "condition"
de.markers <- FindAllMarkers(LungMu_obj_sub, logfc.threshold = 0.25,min.pct = 0.1,only.pos = TRUE)

LungMu_obj_sub <- subset(LungMu_obj, label == "Arterial/Cap pulmonary EC")
Idents(LungMu_obj_sub) <- "condition"
de.markers <- FindAllMarkers(LungMu_obj_sub, logfc.threshold = 0.25,min.pct = 0.1,only.pos = TRUE)
```

## Compute gene per condition in each cell type
```{r compute gene per condition in each cell type} 
analysis_type <- "marker_genes_all_among_conditions"
type <- "All_cells"
Selgenes <- c(
              "Epb41", "Mcf2l","Mcam","Glp1r", "Sox11", "Samd5", "Vwf", "Cd93",
              "Plod1", "Rin3", "Tmx3", "Rock2", "Csrnp1", "Sema6a", "Mapt", "Arap2", "Fmo1", "Meis1", "Gnaq",
              "Muc5b","Pim3", "Depp1","Nrarp","Tob2", "Nupr1", "Nectin2",
              "Rassf9",  "Reep6", "Cstf3", "Utp11","Fdx1", "Aldh3a2",
              "Tpcn2", "Bcar3",  "Pxdc1", "Akap5", "Agrn","Nt5e","Prkag3","Ager", "Dlg4", "Gprc5a", "Ankrd1",
              "Ctsh",  "Lrrk2", "Pmvk", "Slc34a2",  "Elovl1", "Lpcat1",
              "Hsph1", "Ahsa2", "Tuba1a","Mt1",
              "Adamtsl2", "Dpep1","Rgs2", "Plekha6","Atp1a2", "Slc7a10", "Inmt",
              "Tagln","Myh11",  "Pdgfrb", "Carmn",
              "Timp2","Cys1","Gas6", "Gbp3" ,
              "Fkbp5", "Tfec", "Ppt1", "Ap3d1", "Fcho2","Fuca1", "Chil3",  "Vim", "Mmp19", "Cebpb","Marco","Ddb1","Sgk1", "Colgalt1","Galnt3","Cd44", "Cotl1","Plk3",
              "Rell1", "Ms4a6c", "Tbc1d8","Cd83", "Cd226","Got1","Pmaip1","Ighm","Stap1"
             )

Idents(LungMu_obj) <- LungMu_obj$condition
Mu_CD <- subset(LungMu_obj, condition  == "Control" )
Mu_HFD <- subset(LungMu_obj, condition  == "PH-cHFpEF")
 
avg_matrix_CD<-compute_avg_gene_expression_matrix(Mu_CD,Selgenes,names(colLab_reorder))
avg_matrix_HFD<-compute_avg_gene_expression_matrix(Mu_HFD,Selgenes,names(colLab_reorder))
```

## Plots gene per condition in each cell type
```{r gene per condition in each cell type, fig.height = 16, fig.width = 9}
avgHeatmap2(Selgenes,avg_matrix_CD, avg_matrix_HFD,
                      colVecIdent = colLab_reorder,colVecCond = colCond,
                      gapVecR=NULL, gapVecC=NULL,cc=T,
                      cr=F, condCol=F,
                      type = type, folder_name =GeneExprDir)
```


## Save annotated object
```{r save annotated object}
saveRDS(LungMu_obj, file= paste0(output_obj_dir, "340821_Mu_Lung_snRNAseq_merged_annotated.rds" ))
```

## Session Information
````{r Session Information}
sessionInfo()
````



